High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-11-09 doc/llr-pldi2013/interface.tex
monkeyed with figures in section 2 - alot

Our interface organizes functionality into objects.  We describe
these objects in five parts: {\em events} for composing operations 
(Section~\ref{subsec:events}), {\em processors} for parallel computation 
(\ref{subsec:procs}), {\em deferred locks} for synchronization
(\ref{subsec:locks}), {\em physical regions} for data layout and movement
(\ref{subsec:phyreg}), and a {\em machine} object for introspection of the underlying
hardware (\ref{subsec:machmodel}).  In all cases (except for the
machine object which is a static singleton), instances of the objects
are light-weight {\em handles} that provide a unique name for an underlying
implementation.
Every handle is valid everywhere in the system, allowing handles to be freely copied,
passed as arguments to other tasks, or stored in the heap
without having to reason about the distributed nature of the system.
%%   Multiple copies of handles are permitted and handles can be 
%% passed by value.  Furthermore, every handle is valid everywhere in the system.  
%% This property allows a client to pass handles by value between computations 
%% without having to reason about the distributed nature of the system.
%We describe how we support this property in more detail in 
%Section~\ref{sec:impl}.

@@

\subsection{Events}
\label{subsec:events}
Events are the primary mechanism for describing dependences between operations in our system.
Listing~\ref{lst:eventapi} shows the interface for events.  An instance of the {\tt Event} type 
names a unique event in the system.  {\tt NO\_EVENT} (line 5) is a special instance
of an event that by definition has always triggered.  The event interface
supports testing whether an event has triggered (line 7) and waiting on
an event to trigger (line 9).  While these methods can be useful, the encouraged
use of events is passing them as preconditions for other operations.  The client can use the
{\tt merge\_events} call (lines 11-12) to create an event that triggers only when every
event in a given set of events has triggered.  An example using events follows in
Section~\ref{subsec:procs}.
%corresponding to the conjunction of a set of events.

In most cases events are created as the result of other
operations and the implementation is responsible for triggering these events.  Clients
can also create a {\tt UserEvent} (line 15) that is triggered explicitly by the client.

The common synchronization pattern in which multiple consumers are dependent on
a set of producers is supported by a {\tt Barrier} object, which is similar
to a user event, but does not trigger until the expected number of {\tt arrive}
operations have occurred.  To better support
composability and asynchronous operation, our interface differs from barriers in other programming
models\cite{MPI} in three fundamental ways:
%The common synchronization pattern in which multiple consumers are dependent on a 
%(potentially different) set of producers is supported by a {\tt Barrier} object,
%which is similar to a {\tt UserEvent}, but does not trigger until the expected
%number of {\tt arrive} operations have occurred.  Barrier constructs exist in other
%programming models\cite{MPI}, but in addition to the decoupling of the waiters from
%the arrivers, our barrier interface has two further improvements to better support
%composability and asynchronous operation:
\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item The set of producers of arrival calls can be different from the set of consumers waiting on the barrier.
\item An arrival operation can be made dependent on another event, allowing a client to
express that an operation should be completed before the barrier can trigger.
\item The expected arrival count of a barrier can be dynamically modified, allowing
barrier creation before needing to know the exact number of arrivals.
\end{itemize}

Both user events and barriers are a sub-type of events and can therefore
be used transparently to express dependences wherever events are used.
%Although user events and barriers are created and triggered differently,
%operations that express dependences on Events can transparently use
%UserEvents or Barriers as well.

@@

\begin{lstlisting}[float={t},label={lst:eventapi},caption={Event Interface.}]
class Event {
  const unsigned id;
  const unsigned gen;
  static const Event NO_EVENT;
  // check if an event has triggered
  bool has_triggered() const;
  // wait on the event
  void wait() const;
  // Merge events together to create a new event
  static Event merge_events(Event ev1, Event ev2);
  static Event merge_events(const set<Event> &to_merge);
};

class UserEvent : public Event {
  static UserEvent create_user_event();
  void trigger() const;
};

class Barrier : public Event {
  static Barrier create_barrier(unsigned expected_arrivals);
  void alter_arrival_count(int delta) const;
  void arrive(unsigned count = 1,
                  Event wait_for = Event::NO_EVENT) const;
};
\end{lstlisting}

\subsection{Processors}
\label{subsec:procs}
Listing~\ref{lst:procapi} shows the interface for {\tt Processor} objects which allow for
the creation of parallel computations.  Processors provide a way of naming 
every computational unit in the machine.  We describe how to discover processor handles
in Section~\ref{subsec:machmodel}.  In the case of our current interface 
processors are either individual cores or discrete GPUs, but this is easily modified 
to include new processor types (line 10).   Processors support a single {\tt spawn}
method (lines 14-15) that launches a new {\em task} on that processor.
Note that because the spawn operation
is invoked on a processor handle, it is possible to launch a task on any
processor from anywhere in the system.  
The spawn call takes an optional event that must trigger before the task can begin.  
The spawn call returns an event that will trigger when the task completes.

Lines 18-24 of Listing~\ref{lst:procapi} show a simple example using 
the processor and event interface.  The {\tt diamond\_task} launches four
sub-tasks with a diamond dependence pattern, as shown in Figure~\ref{fig:procevents}.
The {\tt merge\_events} call
merges the events {\tt eb} and {\tt ec} to create the shared dependence
for the last sub-task.

\begin{lstlisting}[float={t},label={lst:procapi},caption={Processor Interface.},belowskip=0pt]
class Processor {
  const unsigned id;
  typedef unsigned TaskFuncID;
  typedef void (*TaskFuncPtr)
            (const void *args,size_t arglen,Processor p);
  typedef map<TaskFuncID, TaskFuncPtr> TaskIDTable;

  enum Kind {
    CPU_PROC,GPU_PROC,// ... future processor types
  };
  Kind kind() const;

  Event spawn(TaskFuncID func_id,const void *args,size_t arglen,
              Event wait_on = Event::NO_EVENT) const;
};
\end{lstlisting}

@@

@@

\begin{lstlisting}[float={t},label={lst:lockapi},caption={Deferred Lock Interface.}]
class Lock {
  const unsigned id;
  Event lock(Event wait_on = Event::NO_EVENT) const;
  void unlock(Event wait_on = Event::NO_EVENT) const;
  // Create a new lock, destroy an existing lock
  static Lock create_lock(size_t payload_size = 0);
  void *payload_ptr();
  void destroy_lock();
};
\end{lstlisting}

\begin{figure}
\begin{lrbox}{\mylistingbox}
\begin{lstlisting}
void launcher_task(const void *args, size_t arglen, Processor p) {
  // Unpack lock, input events from arguments
  Lock needed = ...
  // Acquire lock, launch first subtask, release lock
  Event lock_event_1 = needed.lock(prev_event_1);
  Event task_event_1 = p.spawn(SUB, args_1, sizeof(args_1), lock_event_1);
  needed.unlock(task_event_1);
  // same for second subtask
  Event lock_event_2 = needed.lock(prev_event_2);
  Event task_event_2 = p.spawn(SUB, NULL, 0, lock_event_2);
  needed.unlock(task_event_2);
}
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}}\\

\centering
\scalebox{0.8}{
\begin{tikzpicture}
  \draw (1.0,1.5) rectangle (6.2,2.4);
  \node (l1) at (3.6,1.95) {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};

  \draw (6.7,1.5) rectangle (8.6,2.4);
  \node (t1) at (7.65,1.95) {\begin{tabular}{c}{\bf run}\\SUB(args\_1)\end{tabular}};

  \draw (9.1,1.5) rectangle (10.5,2.4);
  \node (u1) at (9.8,1.95) {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0.0,1.95) to (1.0,1.95);
  \draw [->] (6.2,1.95) to (6.7,1.95);
  \draw [->] (8.6,1.95) to (9.1,1.95);

  \draw (0.3,0.1) rectangle (1.7,1.0);
  \node (l2) at (1.0,0.55) {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};

  \draw (2.2,0.1) rectangle (4.1,1.0);
  \node (t2) at (3.15,0.55) {\begin{tabular}{c}{\bf run}\\SUB(args\_2)\end{tabular}};

  \draw (4.6,0.1) rectangle (6.0,1.0);
  \node (u2) at (5.3,0.55) {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0.0,0.55) to (0.3,0.55);
  \draw [->] (1.7,0.55) to (2.2,0.55);
  \draw [->] (4.1,0.55) to (4.6,0.55);

  \draw [->,dotted,thick] (0.2,2.5) to (0.5,1.0);
  \draw [->,dotted,thick] (5.8,1.0) to (6.0,1.5);
  \draw [->,dotted,thick] (10.3,1.5) to (10.6,0.0);

@@

Listing~\ref{lst:lockapi} shows the 
interface for locks.  Unlike blocking locks, the {\tt lock} method (line 4) doesn't
block but instead always returns immediately with an event that will be triggered
when the lock has been acquired.  Just like the spawn method, the lock method also 
takes an optional event parameter as a precondition.  Similarly, the {\tt unlock} 
method (line 5) takes an optional event precondition parameter.

An important difference between deferred locks and blocking locks is that the processor
that requests the lock doesn't have to be the one that uses it.  A common
convention in writing code with deferred locks is to acquire locks on behalf of a task being
launched.  Listing~\ref{lst:lockapi} illustrates this with a simple example.  The
{\tt launcher\_task} is calling a sub-task that is going to require the {\tt needed}
lock.  A lock request is issued and the resulting {\tt lock\_event} is used as
a precondition for launching the {\tt SUB} task.  The unlock operation is then
made dependent
%precondition
on the event corresponding to {\tt SUB} task's completion.  The
{\tt SUB} task can run safely while holding the lock and the lock is released when the
{\tt SUB} task completes.  Figure~\ref{fig:lockevents} shows the event dependence graph, with
dotted lines representing a lock operation's dependence on a previous unlock.
Using deferred locks in this style prevents 
compute resources from waiting on locks by ensuring tasks only run when their needed
locks have already been acquired.

When a lock is used to mediate access to a piece of data, the first thing a task will usually
do after being granted the lock is access the data.  To be able to hide the latency of that
access as well, an additional feature of locks in our interface is the association of a lock with a
{\em payload}, a small (i.e. less than 4KB) 
piece of data that is moved around with the lock and guaranteed to be coherent while 
the lock is held.  The size of the payload is specified when the lock is created (line 7) and
the pointer to the local copy of the payload can be obtained from the {\tt payload\_ptr} method
(line 8).

@@

\begin{lstlisting}[float={t},label={lst:regionapi},caption={Physical Region Interface and Example.}]
class Memory {
  const unsigned id;
  size_t size() const;
};

class RegionMetaData {
  const unsigned id;
  static const RegionMetaData NO_REGION;
  // Create and destroy metadatas
  static RegionMetaData create_region(size_t num_elmts, 
                                      size_t elmt_size);
  void destroy_region() const;
  // Create and destroy instances
  RegionInstance create_instance(Memory memory) const;
  void destroy_instance(RegionInstance instance) const;
  // Create a reduction-only instance
  RegionInstance create_instance(Memory memory, 
                        ReductionOpID redopid) const;
};

class RegionInstance {
  const unsigned id;
  static const RegionInstance NO_INST;
  // Copy between instances
  Event copy_to(RegionInstance target, 
                Event wait_on = Event::NO_EVENT);
  Event reduce_to(RegionInstance target, ReductionOpID redopid,
                Event wait_on = Event::NO_EVENT);
};
\end{lstlisting}

@@

Listing~\ref{lst:regionapi} shows a subset of the interface for physical regions.  
Each memory in the machine's memory hierarchy is named by a {\tt Memory} object (line 1).
The next object is a {\tt RegionMetaData} object (line 7), which defines a class of
regions.  Physical regions that belong to that class are created with the {\tt create\_instance}
method (line 16) and represented by {\tt RegionInstance} objects (line 23).  When creating a physical
region the client specifies the memory in which the physical region is going
to be allocated.  To maintain performance transparency, a new physical region can only be allocated
if the memory has sufficient capacity.
There is no automatic coherence of data between physical regions in the
same class.  It is the client's responsibility to manage data coherence using copies
between physical regions (line 28-29).  Like other operations copies can be
predicated on an event and return an event corresponding to completion.

If the only operations a task will perform on a region are reductions, a special reduction-only
instance can be created.  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance, and applied in bulk to 
the target using the {\em reduce\_to} operation.  This has several benefits.  First, it results
in much better performance, as we will show in section~\ref{subsec:reducmicro}), by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Second, a reduction-only instance is often smaller that the target instance, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

@@

An example using reduction-only regions is shown on lines 41-55 of Listing~\ref{lst:regionapi}.
Two reduction-only regions, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 46-47).  (Reduction operations are registered at startup with integer IDs
as function pointers may not be portable across nodes.)  A physical region of the same class {\tt inst3}
is also created in a different memory (line 48).  Two tasks are launched that will apply
reductions locally into their respective reduction buffers (lines 49-50).  The resulting
events are then used to chain reduction copies back to {\tt inst3} after the tasks complete (lines 51-52).
The reduction operations accumulated in each of {\tt inst1} and {\tt inst2} are applied atomically
to {\tt inst3}.  The order in which the reduction operations are performed will be different from
the original order, but the requirement that reduction operations be commutative ensures the final
result is correct.  Finally, a merge on the events from the two reduction copies is 
performed and a final task using the results in {\tt inst3} is launched (lines 53-54).  This event
graph is shown in Figure~\ref{fig:reducevents}.

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is shown in Listing~\ref{lst:machineapi} and allows the
initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (lines 9-10).
Any running task can get a pointer to the machine object with the
{\tt get\_machine} method (line 12) and use it to obtain lists of all the {\tt Memory} (line 13) or 
{\tt Processor} (line 14) handles in the system.  Using the affinity methods (lines 16-19),
the application can also determine which memories are accessible by each
processor and with what performance, as well as which pairs of memories support copy operations.
%are able to efficiently copy data between themselves.

@@

\begin{lstlisting}[float={t},label={lst:machineapi},caption={Machine Interface.}]
class Machine {
  Machine(int *argc, char ***argv,
          const Processor::TaskIDTable &task_table);
  enum RunStyle {
    ONE_TASK_ONLY,ONE_TASK_PER_NODE,
    ONE_TASK_PER_PROCESSOR,
  };
  void run(Processor::TaskFuncID task_id = 0, 
        RunStyle = ONE_TASK_ONLY, const void *args, size_t arglen);

  int get_proc_mem_affinity(vector<ProcMemAffinity> &result,
          Processor restrict_proc = 0, Memory restrict_mem = 0);
  int get_mem_mem_affinity(vector<MemMemAffinity> &result,
          Memory restrict_mem1 = 0, Memory restrict_mem2 = 0);
};
\end{lstlisting}

>>>>

Our interface is shown in Figure~\ref{fig:runtimeapi} and organizes functionality into objects.  We describe
these objects in six parts: {\em events} for composing operations 
(Section~\ref{subsec:events}), {\em processors} for parallel computation 
(\ref{subsec:procs}), {\em barriers} for producer/consumer coordination (\ref{subsec:barriers}),
{\em deferred locks} for more general synchronization
(\ref{subsec:locks}), {\em physical regions} for data layout and movement
(\ref{subsec:phyreg}), and a {\em machine} object for introspection of the underlying
hardware (\ref{subsec:machmodel}).  In all cases (except for the
machine object which is a static singleton), instances of the objects
are light-weight {\em handles} that provide a unique name for an underlying
implementation object.
Every handle is valid everywhere in the system, allowing handles to be freely copied,
passed as arguments to other tasks, or stored in the heap
without having to reason about the distributed nature of the system.
%%   Multiple copies of handles are permitted and handles can be 
%% passed by value.  Furthermore, every handle is valid everywhere in the system.  
%% This property allows a client to pass handles by value between computations 
%% without having to reason about the distributed nature of the system.
%We describe how we support this property in more detail in 
%Section~\ref{sec:impl}.

@@

\subsection{Events}
\label{subsec:events}
Events are the primary mechanism for describing dependences between operations in our system.
Lines 1-13 of Figure~\ref{fig:runtimeapi} show the interface for events.  An instance of the {\tt Event} type 
names a unique event in the system.  {\tt NO\_EVENT} (line 3)
is a special instance
of an event that by definition has always triggered.  The event interface
supports testing whether an event has triggered (line 5) and waiting on
an event to trigger (line 6).  While these methods can be useful, the preferred
use of events is passing them as preconditions for other operations.  The client can use the
{\tt merge\_events} call (line 7) to create an event that triggers only when every
event in a given set of events has triggered.  A code example using events follows in
Section~\ref{subsec:procs}.
%corresponding to the conjunction of a set of events.

In most cases events are created as the result of other
operations and the implementation is responsible for triggering these events.  Clients
can also create a {\tt UserEvent} (line 10) that is triggered explicitly by the client,
and may be used as precondition like any other event.

%Although user events and barriers are created and triggered differently,
%operations that express dependences on Events can transparently use
%UserEvents or Barriers as well.

@@

\begin{figure}
\begin{lrbox}{\mylistingbox}
\begin{lstlisting}
class Event {
  const unsigned id, gen;
  static const Event NO_EVENT;

  bool has_triggered() const;
  void wait() const;
  static Event merge_events(const set<Event> &to_merge);
};

class UserEvent : public Event {
  static UserEvent create_user_event();
  void trigger() const;
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\begin{lrbox}{\mylistingbox}
\begin{lstlisting}[firstnumber=14]
class Processor {
  const unsigned id;
  typedef unsigned TaskFuncID;
  typedef void (*TaskFuncPtr)(void *args,size_t arglen,Processor p);
  typedef map<TaskFuncID, TaskFuncPtr> TaskIDTable;

  enum Kind { CPU_PROC,GPU_PROC /* ... */ };
  Kind kind() const;

  Event spawn(TaskFuncID func_id,const void *args,size_t arglen, Event wait_on) const;
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\begin{lrbox}{\mylistingbox}
\begin{lstlisting}[firstnumber=25]
class Barrier : public Event {
  static Barrier create_barrier(unsigned expected_arrivals);
  void alter_arrival_count(int delta) const;
  void arrive(unsigned count = 1, Event wait_for = NO_EVENT) const;
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\begin{lrbox}{\mylistingbox}
\begin{lstlisting}[firstnumber=30]
class Lock {
  const unsigned id;
  Event lock(Event wait_on = NO_EVENT) const;
  void unlock(Event wait_on = NO_EVENT) const;

  static Lock create_lock(size_t payload_size = 0);
  void *payload_ptr();
  void destroy_lock();
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\begin{lrbox}{\mylistingbox}
\begin{lstlisting}[firstnumber=39]
class Memory {
  const unsigned id;
  size_t size() const;
};

class RegionMetaData {
  const unsigned id;
  static const RegionMetaData NO_REGION;

  static RegionMetaData create_region(size_t num_elmts, size_t elmt_size);
  void destroy_region() const;

  RegionInstance create_instance(Memory memory) const;
  RegionInstance create_instance(Memory memory, ReductionOpID redopid) const;
  void destroy_instance(RegionInstance instance) const;
};

class RegionInstance {
  const unsigned id;

  Event copy_to(RegionInstance target, Event wait_on = NO_EVENT);
  Event reduce_to(RegionInstance target, ReductionOpID redopid, Event wait_on = NO_EVENT);
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\begin{lrbox}{\mylistingbox}
\begin{lstlisting}[firstnumber=62]
class Machine {
  Machine(int *argc, char ***argv, const Processor::TaskIDTable &task_table);

  void run(Processor::TaskFuncID task_id, const void *args, size_t arglen);

  int get_proc_mem_affinity(vector<ProcMemAffinity> &result, ...);
  int get_mem_mem_affinity(vector<MemMemAffinity> &result, ...);
};
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}} \\

\caption{Runtime Interface.\label{fig:runtimeapi}}
\vspace{-4mm}
\end{figure}

\subsection{Processors}
\label{subsec:procs}
Lines 14-24 of Figure~\ref{fig:runtimeapi} show the interface for {\tt Processor} objects which allow for
the creation of parallel computations.  Processors provide a way of naming 
every computational unit in the machine.  We describe how to discover processor handles
in Section~\ref{subsec:machmodel}.  In the case of our current interface 
processors are either individual CPU cores or discrete GPUs, but this is easily modified 
to include new processor types (line 20).   Processors support a single {\tt spawn}
method (line 23) that launches a new {\em task} on that processor.
Note that because the spawn operation
is invoked on a processor handle, it is possible to launch a task on any
processor from anywhere in the system.  
The spawn call takes an optional event that must trigger before the task can begin, and
always returns an event that will trigger when the task completes.

Figure~\ref{fig:procevents} shows a simple example using 
the processor and event interface.  The {\tt diamond\_task} launches four
sub-tasks with a diamond dependence pattern.
The {\tt merge\_events} call
merges the events {\tt eb} and {\tt ec} to create the shared dependence
for the last sub-task.

@@

\subsection{Barriers}
\label{subsec:barriers}
The common synchronization pattern in which multiple consumers are dependent on
a set of producers is supported by a {\tt Barrier} object (lines 25-29 of Figure~\ref{fig:runtimeapi}), which is similar
to a user event, but does not trigger until the expected number of {\tt arrive}
operations have occurred.  To better support
composability and asynchronous operation, our interface differs from barriers in other programming
models\cite{MPI} in three fundamental ways:
%The common synchronization pattern in which multiple consumers are dependent on a 
%(potentially different) set of producers is supported by a {\tt Barrier} object,
%which is similar to a {\tt UserEvent}, but does not trigger until the expected
%number of {\tt arrive} operations have occurred.  Barrier constructs exist in other
%programming models\cite{MPI}, but in addition to the decoupling of the waiters from
%the arrivers, our barrier interface has two further improvements to better support
%composability and asynchronous operation:
\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item The set of producers arriving at the barrier can be different from the set of consumers waiting on the barrier.
\item An arrival operation can be made dependent on another event, allowing a client to
express that an operation should be completed before the barrier can trigger.
\item The expected arrival count of a barrier can be dynamically modified, allowing
producer tasks to be subdivided as necessary, even after the barrier has been created.
\end{itemize}

Like user events, barriers are a sub-type of events and can therefore
be used transparently to express dependences wherever events are used.

@@

\begin{figure}
\begin{lrbox}{\mylistingbox}
\begin{lstlisting}
void lock_example(const void *args, size_t arglen, Processor p) {
  // Unpack lock, input events, target processors from arguments
  Lock needed = ...
  // Acquire lock, launch first subtask, release lock
  Event lock_event_1 = needed.lock(prev_event_1);
  Event task_event_1 = proc1.spawn(SUB, args_1, sizeof(args_1), lock_event_1);
  needed.unlock(task_event_1);
  // same for second subtask
  Event lock_event_2 = needed.lock(prev_event_2);
  Event task_event_2 = proc2.spawn(SUB, args_2, sizeof(args_2), lock_event_2);
  needed.unlock(task_event_2);
}
\end{lstlisting}
\end{lrbox}
\subfigure{\usebox{\mylistingbox}}\\

\centering
\scalebox{0.8}{
\begin{tikzpicture}
  \node (l1) at (2,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};
  \node (t1) at (4.5,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_1)\end{tabular}};
  \node (u1) at (7,1.95) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0,1.95) to (l1.west);
  \draw [->] (l1.east) to (t1.west);
  \draw [->] (t1.east) to (u1.west);

  \node (l2) at (2,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};
  \node (t2) at (4.5,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf run}\\SUB(args\_2)\end{tabular}};
  \node (u2) at (7,0.25) [shape=rectangle,draw] {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  \draw [->] (0,0.25) to (l2.west);
  \draw [->] (l2.east) to (t2.west);
  \draw [->] (t2.east) to (u2.west);

  \draw [->,dotted,thick] (0.5,2.5) to (l2.140);
  \draw [->,dotted,thick] (u2.40) to (l1.220);
  \draw [->,dotted,thick] (u1.320) to (8.5,-0.5);

  %% \draw (1.0,1.5) rectangle (6.2,2.4);
  %% \node (l1) at (3.6,1.95) {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};

  %% \draw (6.7,1.5) rectangle (8.6,2.4);
  %% \node (t1) at (7.65,1.95) {\begin{tabular}{c}{\bf run}\\SUB(args\_1)\end{tabular}};

  %% \draw (9.1,1.5) rectangle (10.5,2.4);
  %% \node (u1) at (9.8,1.95) {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  %% \draw [->] (0.0,1.95) to (1.0,1.95);
  %% \draw [->] (6.2,1.95) to (6.7,1.95);
  %% \draw [->] (8.6,1.95) to (9.1,1.95);

  %% \draw (0.3,0.1) rectangle (1.7,1.0);
  %% \node (l2) at (1.0,0.55) {\begin{tabular}{c}{\bf lock}\\needed\end{tabular}};

  %% \draw (2.2,0.1) rectangle (4.1,1.0);
  %% \node (t2) at (3.15,0.55) {\begin{tabular}{c}{\bf run}\\SUB(args\_2)\end{tabular}};

  %% \draw (4.6,0.1) rectangle (6.0,1.0);
  %% \node (u2) at (5.3,0.55) {\begin{tabular}{c}{\bf unlock}\\needed\end{tabular}};

  %% \draw [->] (0.0,0.55) to (0.3,0.55);
  %% \draw [->] (1.7,0.55) to (2.2,0.55);
  %% \draw [->] (4.1,0.55) to (4.6,0.55);

  %% \draw [->,dotted,thick] (0.2,2.5) to (0.5,1.0);
  %% \draw [->,dotted,thick] (5.8,1.0) to (6.0,1.5);
  %% \draw [->,dotted,thick] (10.3,1.5) to (10.6,0.0);

@@

The interface for deferred locks is shown in lines 30-38 of Figure~\ref{fig:runtimeapi}.
Unlike traditional blocking locks, the {\tt lock} method (line 32) doesn't
wait if the lock is in use but instead always returns immediately with an event that will be triggered
when the lock has been acquired.  Like a processor's spawn method, the lock method also 
takes an optional event parameter as a precondition.  Similarly, the {\tt unlock} 
method (line 33) takes an optional event precondition parameter.

An important difference between deferred locks and blocking locks is that the processor
that requests the lock doesn't have to be the one that uses it.  A common
convention in writing code with deferred locks is to acquire locks on behalf of a task being
launched.  Figure~\ref{fig:lockevents} illustrates this with a simple example.  The {\tt lock\_example}
function needs to launch two subtasks on two othe processors, each with exclusive access to some resource protected by a
lock named {\tt mutex}.  For each subtask, it first requests the lock on the subtaks behalf, making the
subtask wait until the lock has been acquired.  The parent task also requests that the runtime automatically
release the lock after each subtask has completed.  In the event dependency graph, solid arrows represent
the explicit events used in the code, while the dotted line represents the dynamic ordering that results
from the mutual exclusion guarantee of the lock.  (In this example, the second subtask is granted the lock first.)

When a lock is used to mediate access to a piece of data, the first thing a task will usually
do after being granted the lock is access the data.  To be able to hide the latency of that
access as well, an additional feature of locks in our interface is the association of a lock with a
{\em payload}, a small (i.e. less than 4KB) 
piece of data that is moved around with the lock and guaranteed to be coherent while 
the lock is held.  The size of the payload is specified when the lock is created (line 35 of Figure~\ref{fig:runtimeapi}) and
the pointer to the local copy of the payload can be obtained from the {\tt payload\_ptr} method
(line 36).

@@

@@

Lines 39-61 of Figure~\ref{fig:runtimeapi} show a subset of the interface for physical regions.  
Each memory in the machine's memory hierarchy is named by a {\tt Memory} object (line 39).
The next object is a {\tt RegionMetaData} object (line 44), which defines a class of
regions.  Physical regions that belong to that class are created with the {\tt create\_instance}
method (line 51) and represented by {\tt RegionInstance} objects (line 56).  When creating a physical
region the client specifies the memory in which the physical region is going
to be allocated.  To maintain performance transparency, a new physical region can only be allocated
if the memory has sufficient capacity.
There is no automatic coherence of data between physical regions in the
same class.  It is the client's responsibility to manage data coherence using copies
between physical regions (line 59).  Like other operations copies can be
predicated on an event and return an event corresponding to completion.

If the only operations a task will perform on a region are reductions, a special reduction-only
instance can be created (line 52).  Rather than applying each reduction separately to the target instance,
the reduction operations can be accumulated in the reduction instance, and applied in bulk to 
the target using the {\em reduce\_to} operation (line 60).  This has several benefits.  First, it results
in much better performance, as we will show in section~\ref{subsec:reducmicro}), by reducing both
the total inter-node traffic and the overhead of that traffic while still allowing parallelism.
Second, a reduction-only instance is often smaller that the target instance, potentially improving
cache performance and helping address capacity issues for smaller memories (e.g. a GPU's device
memory).  For example, if reductions are being made to a single field in a larger structure, the 
reduction-only instance only needs to store that field and not the other fields that won't be 
modified by the reduction operation.

@@

An example using reduction-only regions is given in Figure~\ref{fig:reducevents} along with the resulting event graph.
Two reduction-only regions, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 7-8).  (Reduction operations are registered at startup with integer IDs
as function pointers may not be portable across nodes.)  A physical region of the same class {\tt inst3}
is also created in a different memory (line 9).  Two tasks are launched that will apply
reductions locally into their respective reduction buffers (lines 10-11).  The resulting
events are then used to chain reduction copies back to {\tt inst3} after the tasks complete (lines 12-13).
The reduction operations accumulated in each of {\tt inst1} and {\tt inst2} are applied atomically
to {\tt inst3}.  The order in which the reduction operations are performed will be different from
the original order, but the requirement that reduction operations be commutative ensures the final
result is correct.  Finally, a merge on the events from the two reduction copies is 
performed and a final task using the results in {\tt inst3} is launched (lines 14-15).

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is the {\tt Machine} object, shown in lines 62-73 of Figure~\ref{fig:runtimeapi},
which provides initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (lines 65).
Any running task can get a pointer to the machine object with the
{\tt get\_machine} method (line 67) and use it to obtain lists of all the {\tt Memory} (line 68) or 
{\tt Processor} (line 69) handles in the system.  Using the affinity methods (lines 71-72),
the application can also determine which memories are accessible by each
processor and with what performance, as well as which pairs of memories support copy operations.
%are able to efficiently copy data between themselves.

@@
